Why is it that statistical tests for residuals are not widely used? An explanation using visual inference.

### 2023-03-01 paper_comparison/paper_comparison/paper_comparison.Rmd
typo
rollback to previous figures

  # Theme
  theme_light(base_size = 5) +
  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.5)) +
  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
  scale_x_continuous(breaks = seq(0, 3, 0.5), minor_breaks = NULL) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  labs(col = "", size = "# lineups")
```

@@

target_sigma <- c(5, 3.03, 1.84, 1.12, 0.68, 0.41, 0.25)

# Calculate the effect size 
es <- map_dbl(target_sigma, 
              function(.x){
                mod <- poly_model(shape = 2, 
                              x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, 
                              sigma = .x)
                log(mod$sample_effect_size(mod$gen(100), type = "simple"))
                }
              )
```

```{r polypower, fig.height = 3, fig.width = 5, fig.cap = "Comparison of power between different tests for non-linear patterns (uniform fitted values only). Main plot shows the power curves estimated using logistic regression, with dots indicating human evaluations of lineups. Surrounding lines of the visual test show the estimated power of 500 bootstrap samples. Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot. Where would you draw the line of too much non-linearity in the residuals? For the RESET test this is around log effect size 1.5, but for the visual test it is around 2.", dev = 'png', dpi = 450}

@@

# Use area() to design layout
map(1:7, ~patchwork::area(21, .x, 22, .x)) %>%
  reduce(c) %>%
  c(patchwork::area(3, 1, 20, 7), .) %>%
  patchwork::wrap_plots(append(plist, list(p), after = 0), 
                        design = .)
```

@@

# The prediction of the visual test
pred_visual_heter <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 1),
                                offset0 = log(0.05/0.95)) %>%
  mutate(power = predict(visual_heter_glmmod, 
                         type = "response",
                         newdata = data.frame(effect_size = effect_size,
                                              offset0 = offset0))) %>%
  mutate(log_effect_size = log(effect_size))

# The prediction of the boot visual test
pred_visual_heter_boot <- map_df(1:500, function(boot_id) {
  slice_sample(visual_heter_dat, n = nrow(visual_heter_dat), replace = TRUE) %>%
    nest(dat = c(effect_size, offset0, reject)) %>%
    mutate(mod = map(dat, 
                     ~glm(reject ~ effect_size - 1, 
                          family = binomial(), 
                          data = .x,
                          offset = offset0))) %>%
    mutate(power = map(mod, function(mod) {
      result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 1),
                           offset0 = log(0.05/0.95))
      result$power <- predict(mod, type = "response", newdata = result)
      result
    })) %>%
    select(-dat, -mod) %>%
    unnest(power) %>%
    mutate(log_effect_size = log(effect_size)) %>%
    mutate(boot_id = boot_id)
  })
```

@@

  # Theme
  theme_light(base_size = 5) +
  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.5)) +
  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
  # scale_size_manual(values = seq(1, 9)) +
  scale_x_continuous(breaks = seq(1, 7, 1), minor_breaks = NULL) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  labs(col = "", size = "# lineups")
```

@@

target_b <- c(0.27, 0.74, 2.01, 5.46, 14.84, 40.34, 109.66)

# Calculate the effect size 
es <- map_dbl(target_b, 
              function(.x){
                mod <- heter_model(a = 0, 
                                   b = .x, 
                                   x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))})
                log(mod$sample_effect_size(mod$gen(100), type = "simple"))
                }
              )
```

```{r heterpower, fig.height = 3, fig.width = 5, fig.cap = "Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only). Main plot shows the power curves, with dots indicating human evaluations of lineups. Surrounding lines of the visual test show the estimated power of 1000 bootstrap samples. Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot. Where would you draw the line of too much heteroskedasticity in the residuals? For the BP test this is around log effect size 3, but for the visual test it is around 5.", dev = 'png', dpi = 450}

# 10091
# 101
# 105
# 106
# 112
# 119
# 133
set.seed(112)

@@

```{r poly-power-uniform-j, fig.height = 4, fig.cap = 'Power of conventional tests and visual tests on lineups containing four different non-linearity shapes. Power curves with higher order of non-linearity are drawn with deeper colours. The default RESET tests under-perform significantly in detecting "Tripe-U" shape. To achieve a similar power as other shapes, a higher order polynomial parameter needs to be used for the RESET test. But this means the order needs to be known prior to testing.'}

@@

# Manually fit the model
dat_heter_a %>%
  nest(dat = c(effect_size, offset0, reject)) %>%
  mutate(mod = map(dat, 
                   ~glm(reject ~ effect_size - 1, 
                        family = binomial(), 
                        data = .x,
                        offset = offset0))) %>%
  mutate(power = map(mod, function(mod) {
    result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 1),
                         offset0 = log(0.05/0.95))
    result$power <- predict(mod, type = "response", newdata = result)
    result
  })) %>%
  select(-dat, -mod) %>%
  unnest(power) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_line(aes(log_effect_size, power, col = a), size = 1) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  scale_color_manual(values = c(rcartocolor::carto_pal(3, "Earth")[1], "grey", rcartocolor::carto_pal(3, "Earth")[3])) +
  theme_light() +
  theme(legend.position = "bottom") +
  facet_wrap(~conv_or_not) +
  labs(col = "Heteroskedasticity shape")
```

>>>>

  # Theme
  theme_light(base_size = 5) +
  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.5)) +
  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
  scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  labs(col = "", size = "# lineups")
```

@@

target_sigma <- c(3.65, 2.84, 2.21, 1.72, 1.34, 1.04, 0.81, 0.63, 0.49, 0.38, 0.3)

# Calculate the effect size 
es <- map_dbl(target_sigma, 
              ~log(poly_model(shape = 2, 
                              x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, 
                              sigma = .x)$
                     average_effect_size(n = 100, type = "kl")
                   )
              )
```

```{r polypower, fig.height = 3, fig.width = 5, fig.cap = "Comparison of power between different tests for non-linear patterns (uniform fitted values only). Main plot shows the power curves estimated using logistic regression, with dots indicating human evaluations of lineups. Surrounding lines of the visual test show the estimated power of 500 bootstrap samples. Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot. Where would you draw the line of too much non-linearity in the residuals? For the RESET test this is around log effect size 2.5, but for the visual test it is around 3.5.", dev = 'png', dpi = 450}

@@

# Use area() to design layout
map(1:11, ~patchwork::area(21, .x, 22, .x)) %>%
  reduce(c) %>%
  c(patchwork::area(3, 1, 20, 11), .) %>%
  patchwork::wrap_plots(append(plist, list(p), after = 0), 
                        design = .)
```

@@

# The prediction of the visual test
pred_visual_heter <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
                                offset0 = log(0.05/0.95)) %>%
  mutate(power = predict(visual_heter_glmmod, 
                         type = "response",
                         newdata = data.frame(effect_size = effect_size,
                                              offset0 = offset0))) %>%
  mutate(log_effect_size = log(effect_size))

# The prediction of the boot visual test
pred_visual_heter_boot <- map_df(1:500, function(boot_id) {
  slice_sample(visual_heter_dat, n = nrow(visual_heter_dat), replace = TRUE) %>%
    nest(dat = c(effect_size, offset0, reject)) %>%
    mutate(mod = map(dat, 
                     ~glm(reject ~ effect_size - 1, 
                          family = binomial(), 
                          data = .x,
                          offset = offset0))) %>%
    mutate(power = map(mod, function(mod) {
      result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
                           offset0 = log(0.05/0.95))
      result$power <- predict(mod, type = "response", newdata = result)
      result
    })) %>%
    select(-dat, -mod) %>%
    unnest(power) %>%
    mutate(log_effect_size = log(effect_size)) %>%
    mutate(boot_id = boot_id)
  })
```

@@

  # Theme
  theme_light(base_size = 5) +
  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.5)) +
  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
  # scale_size_manual(values = seq(1, 9)) +
  scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  labs(col = "", size = "# lineups")
```

@@

target_b <- c(0.17, 0.23, 0.32, 0.46, 0.67, 1.04, 1.7, 3, 6.3, 16.2, 61)

# Calculate the effect size 
es <- map_dbl(target_b, 
              ~log(heter_model(a = 1, 
                               b = .x, 
                               x = {raw_x <- rand_normal(sigma = 0.3);closed_form(~stand_dist(raw_x))})$
                     average_effect_size(n = 100, type = "kl")
                   )
              )
```

```{r heterpower, fig.height = 3, fig.width = 5, fig.cap = "Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only). Main plot shows the power curves, with dots indicating human evaluations of lineups. Surrounding lines of the visual test show the estimated power of 1000 bootstrap samples. Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot. Where would you draw the line of too much heteroskedasticity in the residuals? For the BP test this is around log effect size 2.5, but for the visual test it is around 3.5.", dev = 'png', dpi = 450}

# 10091
# 101
# 105
# 106
# 112
# 119
# 133
set.seed(135)

@@

```{r poly-power-uniform-j, fig.height = 4, fig.cap = 'Power of conventional tests and visual tests on lineups containing four different non-linearity shapes. Power curves with higher order of non-linearity are drawn with deeper colours. The default RESET tests under-perform significantly in detecting "Triple-U" shape. To achieve a similar power as other shapes, a higher order polynomial parameter needs to be used for the RESET test. But this means the order needs to be known prior to testing.'}

@@

# Manually fit the model
dat_heter_a %>%
  nest(dat = c(effect_size, offset0, reject)) %>%
  mutate(mod = map(dat, 
                   ~glm(reject ~ effect_size - 1, 
                        family = binomial(), 
                        data = .x,
                        offset = offset0))) %>%
  mutate(power = map(mod, function(mod) {
    result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
                         offset0 = log(0.05/0.95))
    result$power <- predict(mod, type = "response", newdata = result)
    result
  })) %>%
  select(-dat, -mod) %>%
  unnest(power) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_line(aes(log_effect_size, power, col = a), size = 1) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  scale_color_manual(values = c(rcartocolor::carto_pal(3, "Earth")[1], "grey", rcartocolor::carto_pal(3, "Earth")[3])) +
  theme_light() +
  theme(legend.position = "bottom") +
  facet_wrap(~conv_or_not) +
  labs(col = "Heteroskedasticity shape")
```

### 2023-03-01 paper_comparison/paper_comparison/paper_comparison.tex
typo
rollback to previous figures

\caption{Comparison of power between different tests for non-linear patterns (uniform fitted values only). Main plot shows the power curves estimated using logistic regression, with dots indicating human evaluations of lineups. Surrounding lines of the visual test show the estimated power of 500 bootstrap samples. Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot. Where would you draw the line of too much non-linearity in the residuals? For the RESET test this is around log effect size 1.5, but for the visual test it is around 2.}\label{fig:polypower}
\end{figure}

@@

\caption{Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only). Main plot shows the power curves, with dots indicating human evaluations of lineups. Surrounding lines of the visual test show the estimated power of 1000 bootstrap samples. Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot. Where would you draw the line of too much heteroskedasticity in the residuals? For the BP test this is around log effect size 3, but for the visual test it is around 5.}\label{fig:heterpower}
\end{figure}

@@

\caption{Power of conventional tests and visual tests on lineups containing four different non-linearity shapes. Power curves with higher order of non-linearity are drawn with deeper colours. The default RESET tests under-perform significantly in detecting "Tripe-U" shape. To achieve a similar power as other shapes, a higher order polynomial parameter needs to be used for the RESET test. But this means the order needs to be known prior to testing.}\label{fig:poly-power-uniform-j}
\end{figure}

>>>>

\caption{Comparison of power between different tests for non-linear patterns (uniform fitted values only). Main plot shows the power curves estimated using logistic regression, with dots indicating human evaluations of lineups. Surrounding lines of the visual test show the estimated power of 500 bootstrap samples. Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot. Where would you draw the line of too much non-linearity in the residuals? For the RESET test this is around log effect size 2.5, but for the visual test it is around 3.5.}\label{fig:polypower}
\end{figure}

@@

\caption{Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only). Main plot shows the power curves, with dots indicating human evaluations of lineups. Surrounding lines of the visual test show the estimated power of 1000 bootstrap samples. Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot. Where would you draw the line of too much heteroskedasticity in the residuals? For the BP test this is around log effect size 2.5, but for the visual test it is around 3.5.}\label{fig:heterpower}
\end{figure}

@@

\caption{Power of conventional tests and visual tests on lineups containing four different non-linearity shapes. Power curves with higher order of non-linearity are drawn with deeper colours. The default RESET tests under-perform significantly in detecting "Triple-U" shape. To achieve a similar power as other shapes, a higher order polynomial parameter needs to be used for the RESET test. But this means the order needs to be known prior to testing.}\label{fig:poly-power-uniform-j}
\end{figure}
